Distribution-oblivious Programming with Dish

### 2019-10-03 pldi/draft.tex
Update intro

For a user with many computers and larger inputs, however, developing $p_1$'s distributed equivalent requires a significant effort.
% The scope of such rewrites, and therefore the cost of manual effort, can vary considerably.
Such cost is at best moderate when pipelines can be expressed in domain-specific frameworks---for example, Hadoop and Spark provide a few purely functional primitives that 
As long as the program fits into the model this is only moderately difficult;
For simple, one-off pipelines that take a few minutes to compose this cost---order of magnitude higher than---becomes unjustifiable.

More realistic programs, such as the pipelines presented in this paper, require distributed programming language or a distributed operating system---\eg Plan9's \ttt{rc} shell.
For many legacy pipelines  
Could the generation and execution of $p_1$'s distributing version be fully and correctly automated?

 has to choose between three options, all of which require significant manual effort.
The most invasive is to rewrite the entire program in a distributed programming language
The most popular is to leverage

Use a distributed programming language, even higher rewrites
Or use some form of annotations

 this requires rewriting portions of the program in a new language and use operations

In the simplest case, annotations

Often, they only focus on a few parts of the system---for example, upgrading to a distributed storage layer.
More rarely, companies rewrite entire systems (\eg Twitter's Ruby-to-Scala rewrite~\cite{twitter}), a process that is notoriously difficult under schedule constraints and competitive pressures~\cite{rewrite1, rewrite2}.
The manual effort is expensive, and can introduce new bugs, cascading changes, or regressions of previously fixed performance issues, especially since software today makes extensive use of third-party modules~\cite{breakapp:plos:2017}.

>>>>

For a user with many computers and larger inputs, however, developing $p_1$'s distributed equivalent requires a significant effort.
% The scope of such rewrites, and therefore the cost of manual effort, can vary considerably.
In simple cases comprised of pure functions, has to be rewritten in the language and interfaces exposed by the chosen big-data framework or DSL.
This cost is unjustifiable for one-off pipelines that take a few minutes to compose (but are applied to large datasets).
More complex pipelines, such as the ones presented in later sections, require full-fledged distributed programming language or a distributed operating system---\eg Plan9's \ttt{rc} shell.
In these cases, manual rewriting is expensive and can introduce new bugs, cascading changes, or divergence from legacy functionality---
Could the generation and execution of $p_1$'s distributed version be fully and correctly automated?

% Some (parts of) pipelines require only moderate effort, as they are 
% For some pipelines, this effort is merely moderate, as 
%  can be expressed in domain-specific frameworks---for example, Hadoop and Spark provide a few purely functional primitives that 
%   ---still requiring rewriting the program in a different (\eg Scala for Spark).
% 
% This effort is moderate if the program expressed by these pipelines fits into a domain-specific framework.
% At times, such pipelines
% As long as the program fits into the model this is only moderately difficult;
% 
% For many legacy pipelines  
%  has to choose between three options, all of which require significant manual effort.
% The most invasive is to rewrite the entire program in a distributed programming language
% The most popular is to leverage 
% 
% Use a distributed programming language, even higher rewrites
% Or use some form of annotations 
% 
%  this requires rewriting portions of the program in a new language and use operations
% 
% 
% In the simplest case, annotations
% 
% Often, they only focus on a few parts of the system---for example, upgrading to a distributed storage layer.
% More rarely, companies rewrite entire systems (\eg Twitter's Ruby-to-Scala rewrite~\cite{twitter}), a process that is notoriously difficult under schedule constraints and competitive pressures~\cite{rewrite1, rewrite2}.
%  especially since software today makes extensive use of third-party modules~\cite{breakapp:plos:2017}.
%
